\mythesistitleNOTSYNPOSIS \tnote{Synopsis title is ``Generalization of Consecutive Ones Property of Binary Matrices''}

### 2012-01-14 src/030chap3.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
  SEC 4.}  This chapter documents all the new results obtained by us
in the area of tree path labeling of path hypergraphs which is the
parent problem addressed in this thesis.

@@

\section{Introduction}
\label{sec:myresearchintro}
In Section~\ref{sec:motive} we see that consecutive-ones property and
its equivalent problem of interval labeling of a hypergraph is a
special case of the general problem of tree path labeling of path
hypergraphs.  The problem of consecutive-ones property testing can be
easily seen as a simple constraint satisfaction problem involving a
hypergraph or a system of sets from a universe. Every column (row) of
the binary matrix can be converted into a set of non-negative integers
which are the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to an
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the universe. The cardinality of the interval
assigned to it must be same as the cardinality of the set, and the
intersection cardinality of any two sets must be same as the
intersection cardinality of the corresponding intervals. This is a
necessary and sufficient condition. \tnote{(ICPIA) put reference to
  survey chapter}

@@

To characterize a feasible TPL, we consider the case were a tree path
labeling is also given as input and we are required to test if the
given labeling is feasible. This is defined by the \FTPL problem.

@@

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. Hence we consider two special cases of \CFTPL\ -- namely, \CFTPLINT
and \CFTPLKTREE on special target trees, intervals and \kstars,
respectively.  Section~\ref{sec:spltargettree} discusses these
problems.

@@

An algorithm for \CFTPL on general trees which is less efficient than
polynomial time is presented in
Section~\ref{sec:norestraint}. \tnote{put some more description
  here. refer WG11.}

\remove{We focus on the question of generalizing the notion of an
  ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
  show that for a given set system $\cF$, a tree $T$, and an
  assignment of paths from $T$ to the sets, there is a feasible
  bijection between $U$ and $V(T)$ if and only if all intersection
  cardinalities among any three sets (not necessarily distinct) is
  same as the intersection cardinality of the paths assigned to them
  and the input runs a filtering algorithm (described in this paper)
  successfully.  This characterization is proved constructively and it
  gives a natural data structure that stores all the relevant feasible
  bijections between $U$ and $V(T)$.  Further, the filtering algorithm
  is also an efficient algorithm to test if a tree path labeling to
  the set system is feasible.  This generalizes the result in
  cite-nsnrs09.}

@@

{\bf Hypergraph Preliminaries:} The set $\F \subseteq (2^{U} \setminus
\emptyset)$ is a {\em set system} of a universe $U$ with $|U| = n$.
The {\em support} of a set system $\F$ denoted by $supp(\cF)$ is the
union of all the sets in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For
the purposes of this paper, a set system is required to ``cover'' the
universe; $ supp(\cF) = U$.  A set system $\cF$ can also be visualized
as a {\em hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are
the sets in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a
hypergraph $\cF$ is a graph such that its vertex set has a bijection
to $\cF$ and there exists an edge between two vertices iff their
corresponding hyperedges have a non-empty intersection \cite{mcg04}.

Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
each other, denoted by $\cF' \cong \cF''$, iff there exists a
bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
\cite{kklv10}, written as $B=\phi(A)$.  This is called {\em hypergraph
  isomorphism}.

\add{A set system $\cF$ can be alternatively} represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition
here to path systems. Due to the equivalence of set system and
hypergraph in the scope of this paper, we drop the subscript $_H$ in
the notation and refer to both the structures by $\cF$.

{\bf Path Hypergraph from a Tree:} The graph $T$ represents a {\em
  target tree} with same number of vertices as elements in $U$;
$|V(T)|=|U|=n$.  A {\em path system}\, $\cP$ is a set system of paths
from $T$; $\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{
  is a path} \}$.  This generalizes the fact, from the literature
\cite{bls99,kklv10}, that intervals can be viewed as sub-paths of a
path.

\add{Due to the equivalence of set system and hypergraph in the scope
  of this paper, we drop the subscript $_H$ in the notation and refer
  to both the structures by $\cF$.}

If the intersection graphs of $\cF$ and $\cP$ (a path system) are
isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
  path labeling} of the hypergraph $\cF$.  Note that there are two
kinds of isomorphisms here.  One is the isomorphism of intersection
graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
respectively. Second is the isomorphism between the hypergraphs $\cF$
and $\cP$.

\add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
  the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
  vertex sets that represent the hyperedges for each hypergraph
  respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
  v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
  follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
  emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
  system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
  alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
  most scenarios in this paper, what is given are the pair $(\cF,
  \cl)$ and the target tree $T$; hence this notation will be used more
  often.}

If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
a {\em path hypergraph} and $\cP$ is called {\em path representation}
of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
then it is clear that there is an {\em induced path labeling}
$\cl_\phi: \cF \rightarrow \cP$ to the set system; $\cl_\phi(S) =
\set{y \mid y = \phi(x), x \in S}$ for all $S \in \cF$. Recall that
$supp(\cP) = V(T)$.

\add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
intersection graph $\bI(\cP)$ of a path system $\cP$.  This
isomorphism gives a bijection $\cl': V(G) \rightarrow \cP$. Moreover,
for the purposes of this paper, we require that in a path labeling,
$supp(\cP) = V(T)$.  If graph $G$ is also isomorphic to $\bI(\cF)$ for
some hypergraph $\cF$, then clearly there is a bijection $\cl: \cF
\rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$ where $v_S$ is the
vertex corresponding to set $S$ in $\bI(\cF)$ for any $S \in
\cF$. This bijection $\cl$ is called the {\em path labeling} of the
hypergraph $\cF$ and the path system $\cP$ may be alternatively
denoted as $\cF^\cl$.

A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
\cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this work},
we are given as input $\cF$ and a tree $T$, and the question is
whether there is a path labeling $\cl$ to a set of paths in $T$.  We
refer to such a solution path system by $\cF^\cl$.  A path labeling
$(\cF, \cl)$ is defined to be {\em feasible} if
% $\cF \cong \cF^\cl$ and this
there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
\cF^\cl$ such that $\cl_\phi = \cl$.

{\bf Overlap Graphs and Marginal Hyperedges:} An {\em overlap graph}\,
$\bO(\cF)$ of a hypergraph $\cF$ is a graph such that its vertex set
has a bijection to $\cF$ and there exists an edge between two of its
vertices iff their corresponding hyperedges overlap. Two hyperedges
$S$ and $S'$ are said to {\em overlap}, denoted by $S \overlap S'$, if
they have a non-empty intersection and neither is contained in the
other; $S \overlap S' \text{ iff } S \cap S' \ne \emptyset, S
\nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$ is a spanning
subgraph of $\bI(\cF)$ and not necessarily connected. Each connected
component of $\bO(\cF)$ is called an {\em overlap component}.

A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other marginal hyperedge in $\cF$, then it is called {\em
  super-marginal}.

{\bf $k$-subdivided star -- a special tree} A {\em star} graph is a
complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$ is
the number of leaves. The vertex with maximum degree is called the
{\em center} of the star and the edges are called {\em rays} of the
star. A {\em $k$-subdivided star} is a star with all its rays
subdivided exactly $k$ times. The definition of a {\em ray of a
  $k$-subdivided star} is extended to the path from the center to a
leaf. It is clear that all rays are of length $k+2$.

@@

A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
where second universe is the vertex set $V$ of a given tree $T$ and
every second subset in the ordered pairs is a path in this
tree. Formally, the definition is as follows.
\begin{align*}
  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
    s.t. }T[P_i] \text{ is a path, } i \in I \}
\end{align*}
In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
$\A$. As mentioned before for set systems, the paths cover the whole
tree, i.e. $\bigcup_{i \in I}P_i = V$ \\

\noindent
Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
a set assignment, a path assignment $\A$ is defined to be {\em
  feasible} if there exists a bijection defined as follows.
\begin{align}
  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
  \text{ for all } i \in I, \sigma \text{ is a bijection}
  \label{eq:stf}
\end{align}

Let $X$ be a partially ordered set with $\preccurlyeq$ being the
partial order on $X$.  $mub(X)$ represents an element in $X$ which is
a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
$X$ if $\nexists X_q \in X$ such that $X_m
\preccurlyeq X_q$. \\

@@

In this section we give an algorithmic characterization of a
feasibility of tree path labeling.  Consider a path labeling $(\cF,
\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
  Cardinality Preserving Path Labeling (ICPPL)} if it has the
following properties.

@@

The following lemma is useful in subsequent arguments.
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
  \setminus P_3)|$. Thus lemma is proven. %\qed
\end{proof}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{perms} or {\tt filter\_1}, and prove
its correctness.  This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
  \label{perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\}) \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2
    \setminus S_1 \}$ \label{setbreak} \COMMENT {Remove $S_1$, $S_2$
      and add the ``filtered'' sets} \STATE {\bf for} every $S \in
    \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

@@

    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
    ensured at the end of every iteration.
    % \begin{enumerate} [{Case 3.}1:]
    % \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially true
    %   by induction hypothesis.
    % \item {\em Only one, say $R$, is a new set.}  If $R = S_1 \cap
    %   S_2$, from Lemma~\ref{lem:fourpaths} and invariant III
    %   hypothesis, this case is proven. Similarly if $R$ is any of
    %   the other new sets, the case is proven by also using Lemma
    %   ~\ref{lem:setminuscard}.
    % \item {\em At least two of $R, R', R''$ are new sets.}  The new
    %   sets are disjoint hence this case is vacuously true.
    % \end{enumerate}
  \end{enumerate} %\qed
  % \vspace{-6mm}

>>>>

This chapter documents all the new results obtained by us
in the area of tree path labeling of path hypergraphs which is the
parent problem addressed in this thesis.

@@

\section{Summary of Proposed Problems}
\label{sec:myresearchintro}
In Section~\ref{sec:motive} we see that consecutive-ones property and
its equivalent problem of interval labeling of a hypergraph is a
special case of the general problem of tree path labeling of path
hypergraphs.  The problem of consecutive-ones property testing can be
easily seen as a simple constraint satisfaction problem involving a
hypergraph or a system of sets from a universe. Every column (row) of
the binary matrix can be converted into a set of non-negative integers
which are the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to an
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the universe. The cardinality of the interval
assigned to it must be same as the cardinality of the set, and the
intersection cardinality of any two sets must be same as the
intersection cardinality of the corresponding intervals. This is a
necessary and sufficient condition. \tnote{(ICPIA) put reference to
  survey chapter}

@@

To characterize a feasible TPL, we consider the case where a tree path
labeling is also given as input and we are required to test if the
given labeling is feasible. This is defined by the \FTPL problem.

@@

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. Hence we consider two special cases of the same -- namely, \CFTPLINT
and \CFTPLKTREE on special target trees, namely, intervals and \kstars,
respectively.  Section~\ref{sec:spltargettree} discusses these
problems.

@@

An algorithm for \CFTPL on general trees which is less efficient than
polynomial time is presented in
Section~\ref{sec:norestraint}. \tnote[minor]{put some more description
  here. refer WG11.}

\remove[minor]{We focus on the question of generalizing the notion of an
  ICPIA cite-nsnrs09 to characterize feasible path assignments.  We
  show that for a given set system $\cF$, a tree $T$, and an
  assignment of paths from $T$ to the sets, there is a feasible
  bijection between $U$ and $V(T)$ if and only if all intersection
  cardinalities among any three sets (not necessarily distinct) is
  same as the intersection cardinality of the paths assigned to them
  and the input runs a filtering algorithm (described in this paper)
  successfully.  This characterization is proved constructively and it
  gives a natural data structure that stores all the relevant feasible
  bijections between $U$ and $V(T)$.  Further, the filtering algorithm
  is also an efficient algorithm to test if a tree path labeling to
  the set system is feasible.  This generalizes the result in
  cite-nsnrs09.}

@@

\subsection{Hypergraph Preliminaries}

\begin{definition}[Set systems and hypergraphs] 
  \noindent 
  The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
    system} of a universe $U$ with $|U| = n$.  The {\em support} of a
  set system $\F$ denoted by $\supt{\cF}$\tnote[minor]{change all supp
    references to supt command} is the union of all the sets in $\F$;
  $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this paper, a
  set system is required to ``cover'' the universe; $ supp(\cF) = U$.

  \noindent
  A set system $\cF$ can also be visualized as a {\em hypergraph}
  whose vertex set is $supp(\cF)$ and hyperedges are the sets in
  $\cF$. This is a known representation for
  interval systems in literature \cite{bls99,kklv10}.  We extend this
  definition here to path systems. Due to the equivalence of set
  system and hypergraph in the scope of this paper, we drop the
  subscript $_H$ in the notation and refer to both the structures by
  $\cF$.

  \noindent
  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
  $\cF$ is a graph such that its vertex set has a bijection to $\cF$
  and there exists an edge between two vertices \iff their
  corresponding hyperedges have a non-empty intersection \cite{mcg04}.

  \noindent 
  Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
  each other, denoted by $\cF' \cong \cF''$, \iff there exists a
  bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for
  all sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff
  $B$ is a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
  \cite{kklv10}, written as $B=\phi(A)$.  This is called {\em
    hypergraph isomorphism}.  
\end{definition}

\begin{definition}[Path Hypergraph from a Tree]
  The graph $T$ represents a {\em target tree} with same number of
  vertices as elements in $U$; $|V(T)|=|U|=n$.  A {\em path system}\,
  $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
  \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
  the fact, from the literature \cite{bls99,kklv10}, that intervals
  can be viewed as sub-paths of a path.

  \noindent
  If the intersection graphs of $\cF$ and $\cP$ (a path system) are
  isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
  $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
    path labeling} of the hypergraph $\cF$.\endnote{Note that there
    are two kinds of isomorphisms here.  One is the isomorphism of
    intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and
    $\bI(\cP)$ respectively. Second is the isomorphism between the
    hypergraphs $\cF$ and $\cP$.}

% The subset of the path labeling that
%   maps elements of the universe to paths that contain only a leaf the
%   target tree is  known as a {\em leaf labeling}.
%   \ie if the target tree has the set of $k$ leaves $L = \{v_i, i \in
%   [k]\}$, then 
% $\cl': \rightarrow U' \rightarrow L$ is a leaf labeling where $U'
% \subseteq U$, $|U'| = k$.

%   $\cL_0 = \{ (x_i,v_i) \mid x_i \in
%   U, v_i \in T, x_i \ne x_j, v_i \ne v_j, i \ne j, i,j \in [k] \}$
%   where $k$ is the number of leaves in $T$

  \noindent
  \add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be
    the above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are
    the vertex sets that represent the hyperedges for each hypergraph
    respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
    v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
    follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
    emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the
    path system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$
    may be alternatively denoted in terms of $\cF$ and $\cl$ as
    $\cF^\cl$. In most scenarios in this paper, what is given are the
    pair $(\cF, \cl)$ and the target tree $T$; hence this notation
    will be used more often.}

  \noindent
  If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is
  called a {\em path hypergraph} and $\cP$ is called {\em path
    representation} of $\cF$. If this isomorphism is $\phi: supp(\cF)
  \rightarrow V(T)$, then it is clear that there is an {\em induced
    path labeling} $\cl_\phi: \cF \rightarrow \cP$ to the set system;
  $\cl_\phi(S) = \set{y \mid y = \phi(x), x \in S}$ for all $S \in
  \cF$. Recall that $supp(\cP) = V(T)$.

  \noindent
  \add{A graph} $G$ is a {\em path graph} if it is isomorphic to the
  intersection graph $\bI(\cP)$ of a path system $\cP$.  This
  isomorphism gives a bijection $\cl': V(G) \rightarrow
  \cP$. Moreover, for the purposes of this paper, we require that in a
  path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
  to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
  bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
  where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
  any $S \in \cF$. This bijection $\cl$ is called the {\em path
    labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
  alternatively denoted as $\cF^\cl$.

  \noindent
  A path labeling $(\cF, \cl)$ is defined to be {\em feasible} if $\cF
  \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
  \rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.  \add{In this
    work}, we are given as input $\cF$ and a tree $T$, and the
  question is whether there is a path labeling $\cl$ to a set of paths
  in $T$.  We refer to such a solution path system by $\cF^\cl$.  A
  path labeling $(\cF, \cl)$ is defined to be {\em feasible} if
  % $\cF \cong \cF^\cl$ and this
  there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
  supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF
  \rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.
\end{definition}

\begin{definition}[Overlap Graphs and Marginal Hyperedges]
  An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph
  such that its vertex set has a bijection to $\cF$ and there exists
  an edge between two of its vertices \iff their corresponding
  hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
    overlap}, denoted by $S \overlap S'$, if they have a non-empty
  intersection and neither is contained in the other; $S \overlap S'
  \text{ \iff } S \cap S' \ne \emptyset, S \nsubseteq S', S'
  \nsubseteq S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$
  and not necessarily connected. Each connected component of
  $\bO(\cF)$ is called an {\em overlap component}.

  \noindent
  A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
  \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
  \cite{kklv10}. Additionally, if $S$ is such that it is contained in
  no other marginal hyperedge in $\cF$, then it is called {\em
    super-marginal}.
\end{definition}

\begin{definition}[$k$-subdivided star] 
  A {\em star} graph is a
  complete bipartite graph $K_{1,p}$ which is clearly a tree and $p$
  is the number of leaves. The vertex with maximum degree is called
  the {\em center} of the star and the edges are called {\em rays} of
  the star. A {\em $k$-subdivided star} is a star with all its rays
  subdivided exactly $k$ times. The definition of a {\em ray of a
    $k$-subdivided star} is extended to the path from the center to a
  leaf. It is clear that all rays are of length $k+2$.
\end{definition}

@@

\remove[dire]{
A path assignment }% $\A$ to $\F$ is defined as a set assignment
% where second universe is the vertex set $V$ of a given tree $T$ and
% every second subset in the ordered pairs is a path in this
% tree. Formally, the definition is as follows. 
% \begin{align*}
%   \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
%     s.t. }T[P_i] \text{ is a path, } i \in I \}
% \end{align*}
% In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
% $\A$. As mentioned before for set systems, the paths cover the whole
% tree, i.e. $\bigcup_{i \in I}P_i = V$ \\

% Generalizing the definition of { feasibility} in cite{nsnrs09} to
% a set assignment, a path assignment $\A$ is defined to be {
%   feasible} if there exists a bijection defined as follows.
% \begin{align}
%   \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
%   \text{ for all } i \in I, \sigma \text{ is a bijection}
%   \label{eq:stf}
% \end{align}

\begin{definition}[Partially ordered sets]

Let $X$ be a partially ordered set with $\preccurlyeq$ being the
partial order on $X$.  $mub(X)$ represents an element in $X$ which is
a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
$X$ if $\nexists X_q \in X$ such that $X_m
\preccurlyeq X_q$. 
\end{definition}

@@

\tnote{\sc refer to Problem Defs here}
In this section we give an algorithmic characterization of the
feasibility of a tree path labeling.  Consider a path labeling $(\cF,
\cl)$ on the given tree $T$. We call $(\cF, \cl)$ an {\em Intersection
  Cardinality Preserving Path Labeling (ICPPL)} if it has the
following properties.

@@

The following lemma is useful in subsequent arguments.
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}
\begin{proof}\thesisspacing%[Proof of Lemma~\ref{lem:setminuscard}]
  Let $P_i = \cl(S_i)$, for all $1 \le i \le 3$.  $|S_1 \cap (S_2
  \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1 \cap S_2| -
  |S_1 \cap S_2 \cap S_3|$. Due to properties (ii) and (iii) of ICPPL,
  $|S_1 \cap S_2| - |S_1 \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1
  \cap P_2 \cap P_3| = |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2
  \setminus P_3)|$. %\qed
\end{proof}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty path
hypergraph isomorphism
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler in terms of set
intersections while retaining the set of isomorphisms, if any, between
$\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path labeling is
% feasible, then all intersection cardinalities are preserved,
% i.e. the path labeling is an ICPPL. Algorithm~\ref{al:filterone}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{al:filtertwo} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that the
% sets map to paths, ``preserving'' the path labeling.  The rest of
% the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.
First, we present Algorithm~\ref{al:filterone} which describes {\tt
  filter\_1}, and prove its correctness.  This algorithm refines the
path labeling by processing pairs of paths in $\cF^\cl$ that share a
leaf until no two paths in the new path labeling share any leaf.

\begin{algorithm}[h]
  \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
  \label{al:filterone} % {perms}
  \begin{algorithmic}[\lndisplay]
    \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
    \STATE $j \assign 1$\\
    \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
      $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
      $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
    \{S_1, S_2\}) \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2
    \setminus S_1 \}$ \label{setbreak} \COMMENT {Remove $S_1$, $S_2$
      and add the ``filtered'' sets} \STATE {\bf for} every $S \in
    \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
    S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

@@

    Due to the condition in line~\ref{ln:3waycheck}, this invariant is
    ensured at the end of every iteration.
    % \begin{enumerate} [{Case 3.}1:]
    % \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially true
    %   by induction hypothesis.
    % \item {\em Only one, say $R$, is a new set.}  If $R = S_1 \cap
    %   S_2$, from Lemma~\ref{lem:fourpaths} and invariant III
    %   hypothesis, this case is proven. Similarly if $R$ is any of
    %   the other new sets, the case is proven by also using Lemma
    %   ~\ref{lem:setminuscard}.
    % \item {\em At least two of $R, R', R''$ are new sets.}  The new
    %   sets are disjoint hence this case is vacuously true.
    % \end{enumerate}
  \end{enumerate} %\qed
  % \vspace{-6mm}
\end{proof}
